---
title: "Data exploration"
author: "Alexander Noll"
date: "`r Sys.Date()`"
output: html_document
---

```{r, include = FALSE}
library(dplyr)
library(ggplot2)
```

# Data

## Loading

```{r}
train <- 
    readr::read_csv("~/Desktop/Data_Science_Scale/Winton/Winton/Data/train.csv")
```

## Features

We first want to understand the columns of the data:

* We have 40000 observations of 211 variables

Each observation has 25 features (**Feature_1 - Feature_25**) whose meaning is not further specified, the return from one day ago **Ret_MinusTwo**, the return from the half of the current day **Ret_MinusOne** and the interday return (roughly minutewise) starting from the end of the time period where **Ret_MinusOne** ends. In the training sample, we additionally have the target varialbes **Ret_121 - Ret180**, **Ret_PlusOne** and **Ret_PlusTwo**. Furthermore, there are two weight variables **Weight_Intraday** and **Weight_Daily** needed for evaluation purposes.

## Evaluation

The evaluation function is defined as folows:

```{r}
WMAE <- function(pred_returns, actual_returns, weights){
    # Computes the weighted mean absolute error according to the forumula
    # on the Kaggle page
    #
    # Args: pred_returns: numeric vector of length 62; the first 60
    #          elements correspond to intraday returns, the last two to daily
    #          returns
    #       actual_returns: numeric vector of length 62; analogous to above
    #       weights: numeric vector of length two
    #
    # Returns: weighted mean absolute error
    
    intraday_ret <- 
        sum(abs(pred_returns[1:60] - actual_returns[1:60]) * weights[1])
    daily_ret <- 
        sum(abs(pred_returns[61:62] - actual_returns[61:62]) * weights[2])
    
    return(1 / length(pred_returns) * (intraday_ret + daily_ret))
}
```

## Missing values

Our next step is to understand the missing values in the data frame:

```{r}
PlotMissing <- function(df){
    # Plots number of missing values of a data frame
    #
    # Args: df data frame
    #
    # Returns: Plot with missing values plotted
    
    missing <-
        colSums(is.na(df)) %>%
        matrix(nrow = ncol(df)) %>%
        as.data.frame %>%
        mutate(col_name = colnames(df)) %>%
        rename(n_missing = V1) %>%
        filter(n_missing > 0)
    
    ggplot(missing, aes(x = col_name, y = n_missing)) + 
        geom_bar(stat = "identity") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
}

PlotMissing(train)
```

From the plot it seems that some features look unusable, e.g. features 1 and 10.

## Time series of means:

```{r}
means <-
    data_frame(
        means = 
            train %>% 
            select(starts_with("Ret_")) %>%
            colMeans(na.rm = TRUE),
        sd = 
            train %>%
            select(starts_with("Ret_")) %>%
            apply(2, sd, na.rm = TRUE),
        time = 1 :183
    ) %>%
    mutate(
        means_max = means + 2 * sd,
        means_min = means - 2 * sd
    )

ggplot(means, aes(x = time, y = means)) +
    geom_point() +
    geom_line() + 
    geom_errorbar(aes(x = time, ymax = means_max, ymin = means_min)) +
    scale_y_continuous(limits = c(-0.0001, 0.0001))
        
```

In this plot, nothing really is visible and, in fact, that is to be expected. We can, however, use the just constructed data to look at the autocorrelation coefficient:

```{r}
acf(means$means)
```

Here we can see that there is a quite significant negative autocorrelation at lag one. The fact that it is only at lag one, of course, makes it difficult to predict beyond **Ret_121**.

## Time series of some examples

```{r}
PlotRow <- function(row_num) {
    # Plots time series of a row
    #
    # Args: row_num...integer between and 1 and 40000
    #
    # Returns: plot of time series
    
    time_series <- 
        train[row_num, ] %>% 
        select(starts_with("Ret_")) %>%
        select(-Ret_MinusTwo, -Ret_MinusOne, -Ret_PlusOne, -Ret_PlusTwo) %>%
        as.matrix %>%
        t %>% 
        as.data.frame %>%
        rename(ret = V1) %>%
        mutate(time = 1:179)
    
    ggplot(time_series, aes(x = time, y = ret)) +
        geom_point() + 
        geom_line() + 
        geom_smooth(method = "lm")
}
PlotRow(2)
```