---
title: "Predict One time step"
author: "Alexander Noll"
date: "10. November 2015"
output: html_document
---

```{r, include = FALSE}
library(dplyr)
library(caret)
library(magrittr)
```

As we found out in the document for data exploration, the autocorrelation coefficient at lag 1 is pretty strong (-0.5 for the time series of means, roughly -0.1 for individual time steps). Thus it seems reasonable to try to predict the variable **Ret_121** from **Ret_120** (in the first approach we do not use any of the other variables).

## Data prep

```{r}
set.seed(1)
train <- 
    readr::read_csv("~/Desktop/Data_Science_Scale/Winton/Winton/Data/train.csv")
train_ind <- createDataPartition(train$Ret_121, list = FALSE, p = 0.8)
train_train <- train[train_ind, ]
train_cv <- train[-train_ind, ]
```

We impute missing data by median:

```{r}
impute_model <-
    preProcess(train_train, method = "medianImpute")
train_train %<>% predict(impute_model, .)
train_cv %<>% predict(impute_model, .)
```

## Train model

We use a model that can handle missing data, in order not to have to worry about those at the moment:

```{r}
ret_121_model <-
    train(Ret_121 ~ Ret_120,
          data = train_train,
          method = "lm"
          )

preds <- predict(ret_121_model, newdata = train_cv)
mean(abs(train_cv$Ret_121))
mean(abs(train_cv$Ret_121 - preds))
```

The conclusion is that this model does not work at all.